Attributions for Poverty in Post-Socialist Countries

### 2014-02-20 article2013.Rmd
Alaluvun 5.2. viilaukset

Transfer dependency               Social or private transfers as main source of income                                                                                  lits

Low education                     No or only compulsory level education                                                                                                 lits

@@

Voice and Accountability          indicator of state of democracy. Measures the political process, civil liberties and political rights                                 World Bank
----------------------------

@@

The results above indicate that there are cross-national differences between countries and within country groups which means that there are also some other contextual and individual-level factors which should explain the cross-national differences. Therefore, we will also analyse the importance of country-level economic, social and political conditions and individual-level risk factors related to financial hardship as determinants for attributions for poverty. Because the descriptive results showed that the majority of citizens in the post-socialist countries endorsed social and individual blame types of explanations, the subsequent analyses will focus only on these two explanations.

@@

Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 3, 4 and 5). Correlation coefficients within the post-socialist countries are shown on top of each regression line. Figure 3 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010). In regard to post-socialist countries there is substantial variation between countries in economic change. The trend is very clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.62) and more to individual blame (.71). In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries. However, this is greatly due to economic decline in Great Britain. Hence, without Great Britain we can not see any significant association between economic change and poverty perceptions in Western Europe.

@@

There is also a substantial variation in the level of human development among post-socialist countries. Figure 4 shows that the higher of human development is associated with higher level of social blame (.34). Association holds also in the case of individual blame meaning that the higher the human deveploment is the lower support for support for individual blame explanation (-.48). In this comparison Western countries are better-off in terms of human development and due to substantial variation in dependent variables it is harder to confirm our hypothesis than in the case of post-socialist countries.

@@

In regard to the level of goodness of government we can see major differences are great as expected. Overall association is substantial (.41 for social blame and -.47 for individual blame), and pointing that higher level of democracy is indicating higher support for social blame, whereas lower support for individual blame type of explanation is linked with low figure in good governance indicator. As expected , the variation in the goodness of government in Western Europe is much smaller than in post-socialist countries. In these countries the support for blame types of explanations seems to vanish when the level of democracy increases. Consequently, both social and individual fate types of  explanations gain higher support in more democratic societies.

>>>>

Transfer dependency               Social or private transfers as a main source of income                                                                                  lits

Low education                     No or only compulsory level of education                                                                                                 lits

@@

Voice and Accountability          Indicator of state of democracy. Measures the political process, civil liberties and political rights                                 World Bank
----------------------------

@@

The results above indicate that there are cross-national differences between countries and within country groups which means that there are also some other contextual and individual-level factors which should explain the cross-national differences. Therefore, we will also analyze the importance of country-level economic, social and political conditions and individual-level risk factors related to financial hardship as determinants for attributions for poverty. Because the descriptive results showed that the majority of citizens in the post-socialist countries endorsed social and individual blame types of explanations, the subsequent analyses will focus only on these two explanations.

@@

Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 3, 4 and 5). Correlation coefficients within the post-socialist countries are shown on top of each regression line. Figure 3 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crisis (2007-2010). In regard to post-socialist countries there is substantial variation between countries in economic change. The trend is very clear: the less country has suffered from the financial crisis the less there is support to social blame type of explanation (-.62) and more to individual blame (.71). Thus, the results are in line with previous studies which suggest positive correlation between the economic growth and support for the individualistic perception of the causes of poverty (Gallie and Paugam, 2002; Kallio and Niemelä, 2013). In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries. However, this is greatly due to economic decline in Great Britain. Hence, without Great Britain we can not see any significant association between economic change and poverty perceptions in Western Europe.

@@

In regard to country-level living conditions, we assumed that the high level of human development is associated more with individual than social blame. However, the results indicate that the hypothesis is not supported. Figure 4 shows that higher level of human development is associated with higher level of social blame (.34). Association holds also in the case of individual blame, meaning that the higher the human deveploment is the lower is the support for individual blame explanation (-.48). Western European countries, on the other hand, are more better-off in terms of human development. It seems that there is not any significant association between human development and blame types of explanations in Western European countries (also Kallio and Niemelä, 2013; Lepianka et al. 2010).

@@

Finally, there also large variation between countries in the level of goodness of government. Overall, the association between attributions for poverty and the level of good governance is substantial (.41 for social blame and -.47 for individual blame). Regarding post-socialist countries, however, the direction of association is also here opposite as expected. On the one hand, Figure 5 shows that higher level of democracy is indicating higher support for social blame. On the other hand, individual blame type of explanation is supported more likely in countries with low degree of good governance. In regard to Western European countries, the variation in the goodness of government is much smaller than in post-socialist countries. Also the association between perceptions and the good governance differs. Unlike in post-socialist countries, the support for blame types of explanations seems to vanish when the level of good governance increases.
